Extended Data

### 2021-11-14 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd
add statement about 2 additional "weird" tag years we removed

The raw dendrometer band data were manually inspected before analysis. 
We screened the data for three classes of erros. 
First, when a measurement was drastically different from previous and following measurements, it was assumed to be a human error and the datapoint was removed.
Second, when measurements remained essentially unchanged for several readings, followed by a sudden jump then return to a normal growth pattern, this was assumed to be a case where the band was stuck on the tree bark and then released. In these cases, the full annual record for the tree was removed.
Third, data points that deviated substantially from normal growth patterns, but had unknown causes, were removed. If a majority of the data points fell into this class within a tree-year, the entire year was removed from the analysis.

@@

After fitting the growth model, we removed tree-years with poor fits. 
Models were judged to be poorly fit if certain modeled growth characteristics fell outside of the logical range. 
Modeled fits for tree-years were removed under five conditions:
(1) single day growth rates were $\ge 2$ standard deviations away from the mean for each wood-type (SCBI = 2, Harvard Forest = 34); 
(2) $DOY_{ip}$ was $\ge 2$ standard deviations away from the mean for its xylem architecture group, year, and site (SCBI = 62, Harvard Forest = 106); 
(3) tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.02$ mm; SCBI = 0, Harvard Forest = 66); 
(4) model fit predicted total yearly growth to take longer than 365 days, indicating poor model fit (SCBI = 181, Harvard Forest = 199); 
and (5) models with unexplained sharp spikes in growth rate (SCBI = 0, Harvard Forest = 3). 
At Harvard Forest the tag years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others. 
At SCBI a higher proportion of ring-porous trees were removed, the majority falling under condition 4.

>>>>

The raw dendrometer band data were manually inspected before analysis. 
We screened the data for three classes of errors. 
First, when a measurement was drastically different from previous and following measurements, it was assumed to be a human error and the datapoint was removed.
Second, when measurements remained essentially unchanged for several readings, followed by a sudden jump then return to a normal growth pattern, this was assumed to be a case where the band was stuck on the tree bark and then released. In these cases, the full annual record for the tree was removed.
Third, data points that deviated substantially from normal growth patterns, but had unknown causes, were removed. If a majority of the data points fell into this class within a tree-year, the entire year was removed from the analysis.

@@

After fitting the growth model, we removed tree-years with poor fits. 
Models were judged to be poorly fit if certain modeled growth characteristics fell outside of the logical range. 
Modeled fits for tree-years were removed under five conditions:
(1) single day growth rates were $\ge 2$ standard deviations away from the mean for each wood-type (SCBI = 2, Harvard Forest = 34); 
(2) $DOY_{ip}$ was $\ge 2$ standard deviations away from the mean for its xylem architecture group, year, and site (SCBI = 53, Harvard Forest = 106); 
(3) tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.02$ mm; SCBI = 0, Harvard Forest = 66); 
(4) model fit predicted total yearly growth to take longer than 365 days, indicating poor model fit (SCBI = 150, Harvard Forest = 199); 
and (5) models with unexplained sharp spikes in growth rate (SCBI = 0, Harvard Forest = 3). 
At Harvard Forest the tag years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others. 
An additional two tree-years were removed from SCBI because they were poor fits missed using the previous cleaning criteria. 
At SCBI a higher proportion of ring-porous trees were removed, the majority falling under condition 4.
